### 2021-06-07 Manuscript.Rmd
updated manuscript, added supp data and figures

---
bibliography: references.bib
csl: genomebiology.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
---

@@

@@

```{r summarize-trees, verbose= FALSE, include=F, message = FALSE, echo=FALSE, comment=NA, warning=FALSE}
gene_tree_summary<-list()

#loop the summary over all of the col values of interest
for(i in 1:length(focal_tpm_values)){
gene_tree_summary[[i]]<-summarize_trees(gene_trees_ace_calibrated_pruned,focal_tpm_values[i])
}

#put all of the rows together
gene_tree_summary<-gene_tree_summary %>% dplyr::bind_rows() %>% arrange(gene)

@@

	# Create a tidy data frame of changes in expression
	edges_tidy = edges_ace %>% dplyr::select( gene_tree, S_child, length, node_depth_child, node_age_child, node_age_parent, node_parent, node_child, terminal )
	edges_tidy = cbind( edges_tidy, edges_ace[, grepl("scaled", names(edges_ace))] )
	colnames(edges_tidy) = colnames(edges_tidy) %>% sub( "_scaled", "", .)
	edges_tidy %<>% tidyr::gather( names(edges_tidy)[10:18], key = "treatment", value = "change" )

@@

	# Remove tau for now since it is on a different scale
	edges_tidy %<>% filter(treatment != "tau")

@@

@@

	# Create a tidy data frame of changes in expression
	edges_sim_tidy = edges_sim %>% dplyr::select( gene_tree, S_child, length, node_depth_child, node_age_child, node_age_parent, node_parent, node_child, terminal )
	edges_sim_tidy = cbind( edges_sim_tidy, edges_sim[, grepl("scaled", names(edges_sim))] )
	colnames(edges_sim_tidy) = colnames(edges_sim_tidy) %>% sub( "_scaled", "", .)
	edges_sim_tidy %<>% tidyr::gather( names(edges_sim_tidy)[10:18], key = "treatment", value = "change" )

@@

	# Remove tau for now since it is on a different scale
	edges_sim_tidy %<>% dplyr::filter(treatment != "tau")

@@

edges_tidy_original<-edges_tidy
edges_sim_tidy_original<-edges_sim_tidy
edges_tidy <- edges_tidy[!edges_tidy$treatment=="Bradev",]
edges_sim_tidy <- edges_sim_tidy[!edges_sim_tidy$treatment=="Bradev",]

gene_trees_sim_digests<-unlist(lapply(gene_trees_sim_calibrated_pruned, digest))

change_deciles<-quantile(edges_tidy$change,probs= seq(0, 1, by= 0.1))

linear_model<- edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lm(change~S_child*treatment,data=.)
g_linear_model<- edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% glm(change~S_child*treatment,data=.)
lm_mixed<- edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~S_child*treatment+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE))
lm_mixed_S_null<-edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~S_child+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE), REML=FALSE)
lm_mixed_null<-edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~treatment+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE), REML=FALSE)
lm_mixed_interact_null<-edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~S_child+treatment+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE),REML=FALSE)

#check simulated data

linear_modelsim<-edges_sim_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lm(change~S_child*treatment,data=.)
linear_mixed_modelsim<-edges_sim_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>%lmer(change~S_child*treatment+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE))
lm_mixed_modelsim_S_null<-edges_sim_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~S_child+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE), REML=FALSE)
lm_mixed_modelsim_null<-edges_sim_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~treatment+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE), REML=FALSE)
lm_mixed_modelsim_interact_null<-edges_sim_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% lmer(change~S_child+treatment+(1|gene_tree),data=., control = lmerControl(calc.derivs = FALSE),REML=FALSE)

anova_linear_sim<-anova(linear_modelsim)

#try shuffling values along the branch
xshuffle<-edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% transform(.,change=sample(change))
linear_modelshuffle<-lm(change~S_child*treatment,data=xshuffle)
linear_mixed_modelshuffle<-lmer(change~S_child*treatment+(1|gene_tree),data=xshuffle, control = lmerControl(calc.derivs = FALSE))
linear_mixed_modelshuffle_S_null<-lmer(change~S_child+(1|gene_tree),data=xshuffle, control = lmerControl(calc.derivs = FALSE), REML=FALSE)
linear_mixed_modelshuffle_null<-lmer(change~treatment+(1|gene_tree),data=xshuffle, control = lmerControl(calc.derivs = FALSE), REML=FALSE)
linear_mixed_modelshuffle_interact_null<-lmer(change~S_child+treatment+(1|gene_tree),data=xshuffle, control = lmerControl(calc.derivs = FALSE),REML=FALSE)
anova_linear_modelshuffle<- anova(linear_modelshuffle)

#check if largest changes have an effect, they do not.
#x1<- edges_tidy %>% filter(change>change_deciles[2] & change<change_deciles[10]) %>% dplyr::select(change,gene_tree,S_child,treatment)
#x3<-lm(change~S_child*treatment,data=x1)

anovalm<-anova(linear_model)
#anova(x3)
#S_child, treatment and S_child:treatment are all significant

S_child_marginal<- emmeans(linear_model, pairwise ~ S_child | treatment)

#cld(treatment_marginal,
#alpha=0.05,
#Letters=letters,
#adjust="tukey")

Sum_shuffle = Summarize(change ~ S_child*treatment,  data=xshuffle,digits=3)

Sum_sim= edges_sim_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% Summarize(change ~ S_child*treatment,  data=.,digits=3)

Sum = edges_tidy %>% dplyr::select(change,gene_tree,S_child,treatment) %>% Summarize(change ~ S_child*treatment,  data=.,digits=3)

@@

branch_change<- ggplot(Sum, aes(x = S_new,
                y = mean,
                color = treatment)) +
    geom_errorbar(aes(ymin = mean - sd,
                      ymax = mean + sd),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "plain", size=12)) +
    ylab("Mean change along branch") + xlab("") + ggtitle("Empirical") +theme(plot.title = element_text(hjust = 0.5)) +ylim(c(-15,15)) + scale_color_manual(values=c("#F9B326","#F97F06","#BC98D3","#F4A6AE","#AEC64A","#F7D114","#7ECBE0"))

@@

sim_branch_change<- ggplot(Sum_sim, aes(x = S_new,
                y = mean,
                color = treatment)) +
    geom_errorbar(aes(ymin = mean - sd,
                      ymax = mean + sd),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "plain", size=12)) +
    ylab("") + ggtitle("Simulated") + xlab("") + guides(color=FALSE) + theme(plot.title = element_text(hjust = 0.5)) +ylim(c(-15,15)) + scale_color_manual(values=c("#F9B326","#F97F06","#BC98D3","#F4A6AE","#AEC64A","#F7D114","#7ECBE0"))

@@

var_change_obs_data<-edges_tidy %>% group_by(S_child,treatment) %>% summarise(mean_abs_change=mean(change_abs), var_abs_change=var(change_abs),mean_change=mean(change),var_change=var(change),node_age_parent=node_age_parent[1])

@@

#For all treatments, mean abs change is a function of the node_age of the parent - with greater mean abs change the younger the node age of the parent.

var_change_sim_data<- edges_sim_tidy %>% group_by(S_child,treatment) %>% summarise(mean_abs_change=mean(change_abs), var_abs_change=var(change_abs),mean_change=mean(change),var_change=var(change),node_age_parent=node_age_parent[1])

@@

#this pattern holds with simulated values, suggesting that variance is driven by the internal structure of the tree

@@

@@

```{r covariance, verbose= FALSE, include=F, message = FALSE, echo=FALSE, comment=NA, warning=FALSE}
  edges_cov<-edges_ace_original

	# Only retain edges where both parent and child is a speciation event
	edges_cov %<>% dplyr::filter( (Ev_parent=="S") & (Ev_child=="S") )
	#dim(edges_ace)

	# For each row in edges_ace, identify the expected parent in the species tree for S_child
	expected_parent = phyldog_edges[ match(edges_cov$S_child, phyldog_edges[,2]), 1 ]

	edges_cov %<>% dplyr::filter( (S_parent==expected_parent) )
	#dim(edges_ace)

	# Create a tidy data frame of changes in expression
	edges_tidy_cov = edges_cov %>% dplyr::select( gene_tree, S_child, length, node_depth_child, node_age_child, node_age_parent, node_parent, node_child, terminal )
	edges_tidy_cov = cbind( edges_tidy_cov, edges_cov[, grepl("scaled", names(edges_cov))] )
	colnames(edges_tidy_cov) = colnames(edges_tidy_cov) %>% sub( "_scaled", "", .)
	edges_tidy_cov_superset<-	edges_tidy_cov[,c(1:2,7:8,10:14,16:17)]
	edges_tidy_cov<-edges_tidy_cov[,c(1:2,10:14,16:17)]
	#edges_tidy_cov<-edges_tidy_cov %>% filter(S_child %in% c("9","26","N37","35","N37"))

correlation_branches<-function(child,combn){
  correlation<-function(combn,child){
  u<- edges_tidy_cov %>% dplyr::filter(S_child == child) %>% dplyr::select(combn)
  not_all_na <- function(x) any(!is.na(x))
  u <- u%>% dplyr::select_if(not_all_na) %>% na.omit() 
  if(nrow(u)>100){
  cor_u <- cor(u)
  return(cor_u)
  }
  else{
    return(NA)
  }
}

combn1<-lapply(combn, correlation, child) 
names(combn1)<- lapply(combn,short_pairs_tpm) 
combn1<-combn1[lapply(combn1,length)>0]
combn1<-lapply(combn1,function(x){
  if(length(x)>1){
  x=x[,1]
  x=x[x!=1]
  x<-unname(x)
  return(x)
  }
})
combn1<- combn1 %>% unlist()
return(combn1)
}

focal_tpm_values<- focal_tpm_values[focal_tpm_values!="Bradev"]
pairs_tpm<-combn(focal_tpm_values,2, simplify=FALSE)
S_children<-edges_tidy_cov$S_child %>% unique()

Correlations<-lapply(S_children, correlation_branches,pairs_tpm)
Correlations<-lapply(Correlations,as.data.frame) 
Correlations<-lapply(Correlations,tibble::rownames_to_column)
for(i in 1:length(Correlations)){
  names(Correlations[[i]])<-c("zooid",S_children[[i]])
}
Correlations_new<-Reduce(dplyr::full_join,list(Correlations[[1]],Correlations[[2]],Correlations[[3]],Correlations[[4]],Correlations[[5]],Correlations[[6]],Correlations[[7]],Correlations[[8]],Correlations[[9]],Correlations[[10]],Correlations[[11]],Correlations[[12]]))

rm(Correlations)

Correlations_new<-tidyr::gather(Correlations_new,"S_child","Correlation",-zooid)

Correlations<-Correlations_new %>% tidyr::spread(key=zooid, value=Correlation)

species_tree_annotated<- ape::chronos( phyldog_species_numbered_tree, lambda=1, model="correlated", quiet=TRUE )

tags<-species_tree_annotated %>% as_tibble %>% tibble::add_column(.,S= c(S_child=node_map$S[match(node_map$species,species_tree_annotated$tip.label)], species_tree_annotated$node.label))

tags<-dplyr::left_join(tags,Correlations,by=c("S"="S_child"))

species_tree_annotated<- as.treedata(tags)

trait<-Correlations_new 
trait$S_child<-factor(trait$S_child, levels=species_tree_annotated@data$S)
trait_plot<- trait %>%  filter(S_child%in%c("4","6","9","18","26","32","35"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

trait_plot2<- trait %>%  filter(S_child%in%c("N37"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

trait_plot3<- trait %>%  filter(S_child%in%c("N23"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

trait_plot4<- trait %>%  filter(S_child%in%c("N15"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

trait_plot5<- trait %>%  filter(S_child%in%c("N12"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

trait_plot7<- trait %>%  filter(S_child%in%c("N7"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

trait_plot_all<- trait %>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

p11 <- ggtree(species_tree_annotated) +
  xlim(0,3) +
  geom_tiplab()

###Export all of these plots to build figure

```{r covariance_sim, verbose= FALSE, include=F, message = FALSE, echo=FALSE, comment=NA, warning=FALSE}
  edges_sim_cov<-edges_sim_original

	# Only retain edges where both parent and child is a speciation event
	edges_sim_cov %<>% dplyr::filter( (Ev_parent=="S") & (Ev_child=="S") )
	#dim(edges_ace)

	# For each row in edges_ace, identify the expected parent in the species tree for S_child
	expected_parent = phyldog_edges[ match(edges_sim_cov$S_child, phyldog_edges[,2]), 1 ]

	edges_sim_cov %<>% dplyr::filter( (S_parent==expected_parent) )
	#dim(edges_ace)

	# Create a tidy data frame of changes in expression
	edges_tidy_sim_cov = edges_sim_cov %>% dplyr::select( gene_tree, S_child, length, node_depth_child, node_age_child, node_age_parent, node_parent, node_child, terminal )
	edges_tidy_sim_cov = cbind( edges_tidy_sim_cov, edges_sim_cov[, grepl("scaled", names(edges_sim_cov))] )
	colnames(edges_tidy_sim_cov) = colnames(edges_tidy_sim_cov) %>% sub( "_scaled", "", .)
	edges_tidy_sim_cov_superset<-	edges_tidy_sim_cov[,c(1:2,7:8,10:14,16:17)]
	edges_tidy_sim_cov<-edges_tidy_sim_cov[,c(1:2,10:14,16:17)]

S_children_new<-edges_tidy_sim_cov$S_child %>% unique()

correlation_branches_sim<-function(child,combn){
  correlation<-function(combn,child){
  u<- edges_tidy_sim_cov %>% dplyr::filter(S_child == child) %>% dplyr::select(combn)
  not_all_na <- function(x) any(!is.na(x))
  u <- u%>% dplyr::select_if(not_all_na) %>% na.omit() 
  if(nrow(u)>100){
  cor_u <- cor(u)
  return(cor_u)
  }
  else{
    return(NA)
  }
}

combn1<-lapply(combn, correlation, child) 
names(combn1)<- lapply(combn,short_pairs_tpm) 
combn1<-combn1[lapply(combn1,length)>0]
combn1<-lapply(combn1,function(x){
  if(length(x)>1){
  x=x[,1]
  x=x[x!=1]
  x<-unname(x)
  return(x)
  }
})
combn1<- combn1 %>% unlist()
return(combn1)
}

Correlations_sim<-lapply(S_children_new, correlation_branches_sim,pairs_tpm)
Correlations_sim<-lapply(Correlations_sim,as.data.frame) 
Correlations_sim<-lapply(Correlations_sim,tibble::rownames_to_column)
for(i in 1:length(Correlations_sim)){
  names(Correlations_sim[[i]])<-c("zooid",S_children_new[[i]])
}
Correlations_new_sim<-Reduce(dplyr::full_join,list(Correlations_sim[[1]],Correlations_sim[[2]],Correlations_sim[[3]],Correlations_sim[[4]],Correlations_sim[[5]],Correlations_sim[[6]],Correlations_sim[[7]],Correlations_sim[[8]],Correlations_sim[[9]],Correlations_sim[[10]],Correlations_sim[[11]],Correlations_sim[[12]]))

rm(Correlations_sim)

Correlations_sim_new<-tidyr::gather(Correlations_new_sim,"S_child","Correlation",-zooid)

Correlations_sim<-Correlations_sim_new %>% tidyr::spread(key=zooid, value=Correlation)

species_tree_annotated_sim<- ape::chronos( phyldog_species_numbered_tree, lambda=1, model="correlated", quiet=TRUE )

tags<-species_tree_annotated_sim %>% as_tibble %>% tibble::add_column(.,S= c(S_child=node_map$S[match(node_map$species,species_tree_annotated_sim$tip.label)], species_tree_annotated_sim$node.label))

tags<-dplyr::left_join(tags,Correlations_sim,by=c("S"="S_child"))

species_tree_annotated_sim<- as.treedata(tags)

trait<-Correlations_sim_new 
trait$S_child<-factor(trait$S_child, levels=species_tree_annotated_sim@data$S)
trait_plot<- trait %>%  filter(S_child%in%c("4","6","9","18","26","32","35"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(-1,1))

trait_plot2<- trait %>%  filter(S_child%in%c("N37"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(-1,1))

trait_plot3<- trait %>%  filter(S_child%in%c("N23"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(-1,1))

trait_plot4<- trait %>%  filter(S_child%in%c("N15"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(-1,1))

trait_plot5<- trait %>%  filter(S_child%in%c("N12"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(-1,1))

trait_plot7<- trait %>%  filter(S_child%in%c("N7"))%>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(-1,1))

trait_plot_all<- trait %>% ggplot()+geom_tile(aes(zooid,S_child,fill=Correlation))+theme(axis.text.x=element_text(angle = 30, hjust = 1))+ ggplot2::scale_fill_continuous(low='#BBDEF2', high='#0B4A6D',limits=c(0.5,0.9))

###Export all of these plots to build figure

# Evolution of gene expression across species and specialized zooids in Siphonophora {-}
Catriona Munro^1,2^,  Stefan Siebert^3^, Felipe Zapata^4^, Mark Howison^5^, Casey W. Dunn^6^
\newline
\newline
^1^ Department of Ecology and Evolutionary Biology, Brown University, Providence, RI 02912, USA \newline
^2^ Current address: Collège de France, PSL Research University, CNRS, Inserm, Center for Interdisciplinary Research in Biology, 75005 Paris, France \newline
^3^ XXXXXX Insert Stefan address XXXXXXXXXXX \newline
^4^ Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA \newline
^5^ Watson Institute for International and Public Affairs, Brown University, Providence, RI 02912, USA \newline
^6^ Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA

@@

In this study we use RNA-seq data to conduct differential expression analyses among siphonophore zooids, and one specialized tissue (the pneumatophore, a gas filled float), within siphonophore species, and also map RNA-seq data directly onto gene phylogenies in order to compare gene expression signatures within zooid types across species. We also investigate differential gene expression patterns within species, in particular focusing on three species-specific zooid types that represent lineage specific zooid diversification events.

>>>>

---
bibliography: references.bib
csl: genomebiology.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
    fig_caption: yes
---

@@

rm(list = ls())

@@

@@

	# Create a tidy data frame of changes in expression
	edges_tidy = edges_ace %>% dplyr::select( gene_tree, S_child, length, node_depth_child, node_age_child, node_age_parent, node_parent, node_child, terminal )
	edges_tidy = cbind( edges_tidy, edges_ace[, grepl("scaled", names(edges_ace))] )
	colnames(edges_tidy) = colnames(edges_tidy) %>% sub( "_scaled", "", .)
	edges_tidy %<>% tidyr::gather( names(edges_tidy)[10:21], key = "treatment", value = "change" )

@@

	# Remove tau for now since it is on a different scale
	edges_tidy<-edges_tidy %>% select(-tau)

@@

	#add child values

	child<-edges_ace %>% select(gene_tree, node_parent, node_child, PalmatGasmat_child, GasdevGasmat_child, NecdevGasmat_child, PneGasmat_child)

  edges_tidy<- edges_tidy %>% left_join(child)

  edges_tidy<- edges_tidy %>% left_join(S_new, by=c("S_child"="S"))

@@

	# Create a tidy data frame of changes in expression
	edges_sim_tidy = edges_sim %>% dplyr::select( gene_tree, S_child, length, node_depth_child, node_age_child, node_age_parent, node_parent, node_child, terminal )
	edges_sim_tidy = cbind( edges_sim_tidy, edges_sim[, grepl("scaled", names(edges_sim))] )
	colnames(edges_sim_tidy) = colnames(edges_sim_tidy) %>% sub( "_scaled", "", .)
	edges_sim_tidy %<>% tidyr::gather( names(edges_sim_tidy)[10:21], key = "treatment", value = "change" )

@@

	# Remove tau for now since it is on a different scale
  edges_sim_tidy<-edges_sim_tidy %>% select(-tau)

@@

edges_tidy_original<-edges_tidy
edges_sim_tidy_original<-edges_sim_tidy

Sum_sim= edges_sim_tidy %>% filter(treatment %in% c("PalmatGasmat","GasdevGasmat","NecdevGasmat","PneGasmat")) %>% dplyr::select(change,gene_tree,S_child,treatment) %>% Summarize(change ~ S_child*treatment,  data=.,digits=3)

Sum = edges_tidy %>% filter(treatment %in% c("PalmatGasmat","GasdevGasmat","NecdevGasmat","PneGasmat")) %>% dplyr::select(change,gene_tree,S_child,treatment) %>% Summarize(change ~ S_child*treatment,  data=.,digits=3)

@@

branch_change<- ggplot(Sum, aes(x = S_new,
                y = mean,
                color = treatment)) +
    geom_errorbar(aes(ymin = mean - sd,
                      ymax = mean + sd),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "plain", size=12)) +
    ylab("Mean change along branch") + xlab("") + ggtitle("Empirical") +theme(plot.title = element_text(hjust = 0.5)) +ylim(c(-8,8)) + scale_color_manual(values=c("#F9B326","#AEC64A","#F7D114","#7ECBE0"))

@@

sim_branch_change<- ggplot(Sum_sim, aes(x = S_new,
                y = mean,
                color = treatment)) +
    geom_errorbar(aes(ymin = mean - sd,
                      ymax = mean + sd),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "plain", size=12)) +
    ylab("") + ggtitle("Simulated") + xlab("") + guides(color=FALSE) + theme(plot.title = element_text(hjust = 0.5)) +ylim(c(-8,8)) + scale_color_manual(values=c("#F9B326","#AEC64A","#F7D114","#7ECBE0"))

@@

var_change_obs_data<-edges_tidy %>% filter(treatment %in% c("PalmatGasmat","GasdevGasmat","NecdevGasmat","PneGasmat")) %>% group_by(S_child,treatment) %>% summarise(mean_abs_change=mean(change_abs), var_abs_change=var(change_abs),mean_change=mean(change),var_change=var(change),node_age_parent=node_age_parent[1])

@@

var_change_sim_data<- edges_sim_tidy %>% filter(treatment %in% c("PalmatGasmat","GasdevGasmat","NecdevGasmat","PneGasmat")) %>% group_by(S_child,treatment) %>% summarise(mean_abs_change=mean(change_abs), var_abs_change=var(change_abs),mean_change=mean(change),var_change=var(change),node_age_parent=node_age_parent[1])

@@

@@

### now look at change data w/o ratios

Sum_sim2= edges_sim_tidy %>% filter(treatment %in% c("Palmat", "Gasmat", "Gasdev","Necdev","Pne")) %>% dplyr::select(change,gene_tree,S_child,treatment) %>% Summarize(change ~ S_child*treatment,  data=.,digits=3)

Sum2 = edges_tidy %>% filter(treatment %in% c("Palmat", "Gasmat", "Gasdev","Necdev","Pne")) %>% dplyr::select(change,gene_tree,S_child,treatment) %>% Summarize(change ~ S_child*treatment,  data=.,digits=3)

#add more memorable branch names with letters
Sum2 = data.frame(S_new = S_new$S_new[match(Sum2$S_child, S_new$S)], S_child=Sum2$S_child) %>% dplyr::bind_cols(Sum2, .)

Sum_sim2 = data.frame(S_new = S_new$S_new[match(Sum_sim2$S_child, S_new$S)], S_child=Sum_sim2$S_child) %>% dplyr::bind_cols(Sum_sim2, .)

branch_change2<- ggplot(Sum2, aes(x = S_new,
                y = mean,
                color = treatment)) +
    geom_errorbar(aes(ymin = mean - sd,
                      ymax = mean + sd),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "plain", size=12)) +
    ylab("Mean change along branch") + xlab("") + ggtitle("Empirical") +theme(plot.title = element_text(hjust = 0.5)) +ylim(c(-18,18)) + scale_color_manual(values=c("#F9B326","#F9C153","#AEC64A","#F7D114","#7ECBE0"))

leg2<- get_legend(branch_change2)

branch_change2<-branch_change2 + guides(color=FALSE)

sim_branch_change2<- ggplot(Sum_sim2, aes(x = S_new,
                y = mean,
                color = treatment)) +
    geom_errorbar(aes(ymin = mean - sd,
                      ymax = mean + sd),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "plain", size=12)) +
    ylab("") + ggtitle("Simulated") + xlab("") + guides(color=FALSE) + theme(plot.title = element_text(hjust = 0.5)) +ylim(c(-18,18)) + scale_color_manual(values=c("#F9B326","#F9C153","#AEC64A","#F7D114","#7ECBE0"))

change_figure2<- grid.arrange(tree_reference,branch_change2,sim_branch_change2,leg2, bottom="Branch",widths = c(2,2,0.5),layout_matrix = layout)

var_change_obs_data_noratio<-edges_tidy %>% filter(treatment %in% c("Palmat","Gasdev","Necdev","Pne","Gasmat","Gonmal","Gonfem")) %>% group_by(S_child,treatment) %>% summarise(mean_abs_change=mean(change_abs), var_abs_change=var(change_abs),mean_change=mean(change),var_change=var(change),node_age_parent=node_age_parent[1])

##add human readable branch data
var_change_obs_data_noratio<-data.frame(S_new = S_new$S_new[match(var_change_obs_data_noratio$S_child, S_new$S)], S_child=var_change_obs_data_noratio$S_child) %>% bind_cols(var_change_obs_data_noratio, .)

var_change_obs_noratio<-var_change_obs_data_noratio %>% ggplot()+geom_point(aes(node_age_parent,var_abs_change,col=S_new)) + facet_wrap(~treatment) + xlab(" ") + ylab("")+ggtitle(label="Empirical") + theme(plot.title = element_text(hjust = 0.5))+ guides(col=guide_legend(title="Branch"))

legend2<- get_legend(var_change_obs_noratio)

var_change_obs_noratio<-var_change_obs_noratio + guides(color=FALSE)

var_change_sim_data_noratio<- edges_sim_tidy %>% filter(treatment %in% c("Palmat","Gasdev","Necdev","Pne","Gasmat","Gonmal","Gonfem")) %>% group_by(S_child,treatment) %>% summarise(mean_abs_change=mean(change_abs), var_abs_change=var(change_abs),mean_change=mean(change),var_change=var(change),node_age_parent=node_age_parent[1])

##add human readable branch data
var_change_sim_data_noratio<-data.frame(S_new = S_new$S_new[match(var_change_sim_data_noratio$S_child, S_new$S)], S_child=var_change_sim_data_noratio$S_child) %>% bind_cols(var_change_sim_data_noratio, .)

var_change_sim_noratio<-var_change_sim_data_noratio %>% ggplot()+geom_point(aes(node_age_parent,var_abs_change,col=S_new)) + facet_wrap(~treatment) + xlab(" ") + ylab(" ")+ggtitle(label="Simulated") + theme(plot.title = element_text(hjust = 0.5))+ guides(color=FALSE)

layout3<-rbind(c(1,2),c(3,2))

var_change_plot<- grid.arrange(var_change_obs_noratio,legend2,var_change_sim_noratio, bottom="Node age of parent", left="Variance of change",widths = c(2,0.25),layout_matrix = layout3)

edges_tidy_directional <- edges_tidy %>% filter(treatment%in%c("GasdevGasmat", "PneGasmat", "NecdevGasmat", "PalmatGasmat")) %>% mutate(change_direction= case_when(change<2 & change >-2 ~ paste0("Neutral"), change>=2 ~ paste0("Positive"), change<=-2 ~ paste0("Negative"))) %>% select(gene_tree, S_child, S_new, change, treatment, change_direction, blast_hit)

child2<-edges_ace %>% select(gene_tree, node_parent, node_child, ND_parent, ND_child)

edges_tidy_directional_export <- edges_tidy %>% filter(treatment%in%c("GasdevGasmat", "PneGasmat", "NecdevGasmat", "PalmatGasmat")) %>% mutate(change_direction= case_when(change<2 & change >-2 ~ paste0("Neutral"), change>=2 ~ paste0("Positive"), change<=-2 ~ paste0("Negative"))) %>% left_join(child2) %>%  select(gene_tree, S_child, node_parent, node_child, ND_parent, ND_child, S_new, change, treatment, change_direction, blast_hit)

readr::write_csv(edges_tidy_directional_export, "Supplementary_Data/Supplementaryfile6.csv", na = "NA")

@@

# Evolution of gene expression across species and specialized zooids in Siphonophora {-}
Catriona Munro^1,2^,  Stefan Siebert^3^, Mark Howison^4^, Felipe Zapata^5^, Casey W. Dunn^6^
\newline
\newline
^1^ Department of Ecology and Evolutionary Biology, Brown University, Providence, RI 02912, USA \newline
^2^ Current address: Collège de France, PSL Research University, CNRS, Inserm, Center for Interdisciplinary Research in Biology, 75005 Paris, France \newline
^3^ XXXXXX Insert Stefan address XXXXXXXXXXX \newline
^4^ Research Improving People's Lives (RIPL), Providence, RI, USA \newline
^5^ Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA \newline
^6^ Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA

@@

In this study we use RNA-seq data investigate the functional specialization and evolution of zooids across siphonophore species. To carry out these analyses, we also address three key challenges to study the evolution of gene expression in a comparative framework [@dunn2013phylogenetic]. First, we introduce a new metric to normalize gene expression which is valid for comparison across species. A common metric used to quantify gene expression is Transcripts Per Million (TPM) [@li2011rsem; @wagner2012measurement]. TPM is a relative measure of expression which depends on the number of genes present in a reference. Hence, TPM values are a valid metric when comparing libraries within a single species, but are not directly comparable across species when the references are incomplete, or genes have been gained and lost in the course of evolution. To address this issue, we introduce Transcripts Per Million 10K (TPM10K), a metric which normalizes TPM to account for different sequencing depths among species (see Methods for details).
